Capital Budget Funding in Toronto

### 2024-09-27 paper/paper.qmd
wrote weaknesses and next steps

The unequal distribution of capital funding has significant implications for both present and future generations. Underfunded communities often lack access to essential services such as quality education, healthcare, and public transportation. The National Academies of Sciences, Engineering, and Medicine (2023) emphasize that children growing up in low-income households are more likely to experience intergenerational poverty, which can persist without targeted interventions. Limited access to resources in underfunded wards can hinder social mobility and perpetuate cycles of poverty.

Furthermore, the disparities in funding contribute to a widening wealth gap within the city. Wealthier wards receiving more investment can improve infrastructure, increase property values, and attract further economic development, as discussed by Walks et al. (2016). In contrast, underfunded wards may experience deteriorating infrastructure and services, making them less attractive for investment and leading to economic stagnation. This dynamic exacerbates socioeconomic disparities and can strain social services, as residents in underfunded areas may rely more heavily on social assistance programs due to a lack of access to resources.

@@

### Weaknesses

>>>>

The unequal distribution of capital funding has significant implications for both present and future generations. Underfunded communities often lack access to essential services such as quality education, healthcare, and public transportation. The National Academies of Sciences, Engineering, and Medicine (2023) [REFERENCE] emphasize that children growing up in low-income households are more likely to experience intergenerational poverty, which can persist without targeted interventions. Limited access to resources in underfunded wards can hinder social mobility and perpetuate cycles of poverty.

Furthermore, the disparities in funding contribute to a widening wealth gap within the city. Wealthier wards receiving more investment can improve infrastructure, increase property values, and attract further economic development, as discussed by Walks et al. (2016) [REFERENCE]. In contrast, underfunded wards may experience deteriorating infrastructure and services, making them less attractive for investment and leading to economic stagnation. This dynamic exacerbates socioeconomic disparities and can strain social services, as residents in underfunded areas may rely more heavily on social assistance programs due to a lack of access to resources.

@@

### Weaknesses 
While this study provides interesting results, there are a few limitations. Firstly, the analysis is limited by the focus on median household income, educational attainment, and visible minority populations, potentially missing other important socioeconomic factors like housing costs, employment rates, and age demographics. Next, the reliance on 2021 census data and projected capital budgets through 2031 may not account for recent changes. Lastly, the methodology also focuses on correlations without examining the underlying causes of funding disparities, such as political priorities or historical patterns.

### Next Steps 
To address these limitations, future research should expand the range of socioeconomic variables, such as housing costs and employment rates, to provide a more comprehensive understanding of funding distribution. Longitudinal studies tracking changes in funding and socioeconomic outcomes would help assess the long-term effects of capital investments. Additionally, conducting a comparative study with other cities could reveal best practices for equitable funding.
